Monitoring forest health using hyperspectral imagery: Does feature selection improve the performance of machine-learning techniques?

### 2021-06-07 docs/00-manuscripts/ieee/ieee-manuscript.tex
edited and commented on response to Reviewer 2

Even though the feature-selection step should be solely left to the filter methods in this study, it was ensured a priori to account for features with a pairwise correlation of 1.
Having such features within the data can cause undesired effects during model fitting and feature importance calculation.
Hence, after having calculated all pair-wise correlations between features, for pairs which exceeded the threshold of $1 - 10^{-10}$, the feature with the largest mean absolute correlation across all variables was removed from the dataset.

This preprocessing step reduced the number of covariates to 122 (HR), 86 (VI) and 7467 (NRI).

@@

\subsection{Correlation analysis}

PCA was used for a correlation analysis of the feature sets by inspecting the accumulated percentages of the principal components.
The fewer PCs are needed to reach a high proportion of variance explained the more similar the individual features are to each other.
Results showed that, depending on the feature set, between two (HR) and 42 (NRI) \ac{PC} are needed to explain 95\% of variance.
Hence, a high correlation can be observed for feature set HR which reaches 96\% with only two PCs.
Feature sets NRI and VI are more diverse and require 42 (NRI) and 12 (VI) PCs to explain 95\% of the variance in the data, respectively.

>>>>

Individual features were finally removed from these feature sets if they were numerically equivalent to another feature based on their pairwise correlation being greater than $1 - 10^{-10}$.
This preprocessing step reduced the number of covariates to 122 (HR), 86 (VI) and 7467 (NRI).

@@

\subsection{PCA of feature sets}

PCA was used to assess the complexity of the three feature sets.
Depending on the feature set, 95\% of the variance is explained by two (HR), twelve (VI) and 42 (NRI) \ac{PC}s.
HR features are therefore highly redundant, while the applied feature transformations enrich the data set, at least from an exploratory linear perspective.
